---
title: "DBS Projekt"
author: "Siaw Hui Ngu"
date: "7/2/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
pacman::p_load(tidyverse, ggalt, countrycode, rworldmap, gridExtra, broom, dplyr)
```

## PHASE 2: Datenbereinigung und Datenimport
# a) Import data
```{r cars}
data_file_1 <- file.path("D:/Bioinformatik/Semester4_SS21/Datenbanksysteme/Project", "_suicide_rate.csv")

data_file_2 <- file.path("D:/Bioinformatik/Semester4_SS21/Datenbanksysteme/Project", "gdp.csv")

suicide <- read_csv(data_file_1) %>% 
  select(-c('HDI for year', 'suicides/100k pop')) %>%
  dplyr::rename(gdp_for_year = 'gdp_for_year ($)', 
         gdp_per_capita = 'gdp_per_capita ($)', 
         country_year = 'country-year') %>%
  as.data.frame()

gdp <- read_csv(data_file_2) %>%
  as.data.frame()

suicide
gdp
```

# b) Tidy data:
First, we tidy suicide_rate data by excluding countries with <= 3 years of data. From minimum_years, we know that there are 7 countries which have <= 3 years of data. Thus, 7 countries are removed.
```{r}
suicide <- suicide %>%
  filter(year != 2016) %>% # first exclude 2016 data, because the data from Mongolia is not complete
  select(-country_year)

minimum_years <- suicide %>%
  group_by(country) %>%
  dplyr::summarize(rows = n(), 
            years = rows / 12) %>%
  arrange(years)

suicide <- suicide %>%
  filter(!(country %in% head(minimum_years$country, 7))) 

suicide <- as_tibble(suicide)

suicide
```

Then, we tidy gdp data by excluding years < 1985, because the data in suicide_rate is starting from 1985.
```{r}
gdp <- gdp[ -c(5:29) ]
```


## PHASE 3: Visualisierung des Datensatzes
```{r}
#global_average <- (sum(as.numeric(data$suicides_no)) / sum(as.numeric(data$population))) * 100000

suicide %>%
  group_by(year) %>%
  dplyr::summarize(population = sum(population), 
        suicides = sum(suicides_no), 
        suicides_per_100k = (suicides / population) * 100000) %>%
#  geom_path(data = suicide, mapping = aes(x = year,y = suicides_per_100k), inherit.aes = FALSE ) +
  ggplot(aes(x = "year", y = "suicides_per_100k")) + 
  geom_line(col = "deepskyblue3", size = 1) + 
  geom_point(col = "deepskyblue3", size = 2) + 
#  geom_hline(yintercept = global_average, linetype = 2, color = "grey35", size = 1) +
  labs(title = "Global Suicides (per 100k)",
       subtitle = "Trend over time, 1985 - 2015.",
       x = "Year", 
       y = "Suicides per 100k") +
  scale_x_continuous(breaks = seq(1985, 2015, 2)) + 
  scale_y_continuous(breaks = seq(10, 20))
```

## ------ EXAMPLE OF PLOT ------ 

```{r}
set.seed(2015-04-13)
d = data.frame(x =seq(1,10),
           n = c(0,0,1,2,3,4,4,5,6,6),
           logp = signif(-log10(runif(10)), 2))

par(mar = c(5,5,2,5))
with(d, plot(x, logp, type="l", col="red3", 
             ylab=expression(-log[10](italic(p))),
             ylim=c(0,3)))

par(new = T)
with(d, plot(x, n, pch=16, axes=F, xlab=NA, ylab=NA, cex=1.2))
axis(side = 4)
mtext(side = 4, line = 3, 'Number genes selected')
legend("topleft",
       legend=c(expression(-log[10](italic(p))), "N genes"),
       lty=c(1,0), pch=c(NA, 16), col=c("red3", "black"))
```

